---
title: "Statistical_Computing_with_R_Assignment"
author: "Dagim"
date: "2025-02-19"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1 - Generate a random sample of size 250, using the middle four digits of your ID in set.seed(), from        chi-square distribution with 2 degrees of freedom. Provide descriptive summaries for your data and also justify the reason why chose those summaries. Draw histogram of your data and comment on it.

# Answer

# first lets set or generate the randome variable below
```{r results="hide"}
set.seed(9713)  # my ID: GSR/9713/17  Dagim Alemu
randNums <- rchisq(420, df = 2)  # this generates 420 samples with a degree of freedom of 2 and setting them to a variable called randNums
head(randNums)

```

# basic summary of the generated data above 

```{r results="hide"}
library(e1071)
summary(randNums)  # maximum , min , quartile first and third second being median 🫡
sd(randNums)       # Standard deviation
var(randNums)      # Variance

skewness(randNums) # it measures asymmetry meaning could be left or right skewed 
kurtosis(randNums) # it measures wherether or not the data is picked or not 

# Further justification and reasons Justification:Mean, Median, Min, Max: Give an overview of central tendency and range of data. SD, Variance: Measure the spread of the data. Skewness, Kurtosis: These help analyze the distribution's shape; skewness for asymmetry, kurtosis for tail weight.
```
# Histogram plot 

```{r}
# histogram plot
hist(randNums, breaks = 20, col = "lightblue", border = "darkblue",
     main = "Histogram of Chi-Square Distribution (df=2)",
     xlab = "Score", ylab = "Probability", probability = TRUE)

# Add Chi-Square probability curve for comparison
curve(dchisq(x, df = 2), col = "darkgreen", lwd = 2, add = TRUE)

# Description: This histogram shows how often different scores appear 
# in our sample from a Chi-Square distribution with 2 degrees of freedom. 
# The blue bars represent our data, while the green line shows the expected 
# shape of this distribution.

```
# 2 second question 

# loading stata fiile first 
```{r results="hide"}
library(haven) # should install haven for  package first to load dta for stata files first since the data is using the latest stata file version 
women15_49 <- read_dta('women15_49.dta')
head(women15_49)
```
# setting with my middle id and taking 2500 rows from the data set 
```{r results="hide"}
set.seed(9713)
newDf <- women15_49[sample(nrow(women15_49), 2500), ]
```

# a) Provide socio-demographic(select five of them from the data) table by their knowledge where to get the test.

```{r results="hide"}


#  converting v783 col to factor
newDf$v783 <- factor(newDf$v783, levels = c(0, 1), labels = c("No Clue", "In the Know"))


#  v012:- Respondent's current age  v024:- Region v106:- Highest educational level v130: Religion v151: Sex of household head

# converting them to factors
newDf$v024 <- as.factor(newDf$v024)
newDf$v106 <- as.factor(newDf$v106)
newDf$v130 <- as.factor(newDf$v130)
newDf$v151 <- as.factor(newDf$v151)

# create the table
socio_demographic_table <- table(
  Knows_HIV_Test = newDf$v783,
  Age = newDf$v012,
  Region = newDf$v024,
  Education = newDf$v106,
  Religion = newDf$v130,
  Head_Gender = newDf$v151
)

# Print the table
print(socio_demographic_table)
```

B) Suggest the most appropriate model to assess association between your outcome vari
able and set of your explanatory variables.

```{r results="hide"}
# I choose the logistic regression model because the outcome variable is binary (yes or no)
# Using v012 (age), v024 (region), v106 (education), v130 (religion), and v151 (sex of household head) as explanatory variables

# Fit logistic regression model
model <- glm(v783 ~ v012 + v024 + v106 + v130 + v151, 
             data = newDf, family = binomial(link = "logit"))

```

# c) Using 10 fold cross-validation. Select the model that you think is the best.

# Logistic regression model with no special settings has RMSE of 0.463, R² is 0.088, and MAE is 0.423. This mean model explain little about knowing where to get HIV test. Because R² very low, we can make model much better.

# converted the dta file to csv i had multiple problems with the dta file so i was forced to work with csv file version of the file 

```{r}
# Load libraries
library(tidyverse)
library(caret)

# Load data and set seed
data <- read.csv("women15_49.csv", na.strings = "NA")
set.seed(1234) # Replace with your ID's middle 4 digits
sample_data <- data %>% sample_n(2500)

# Clean outcome variable (v756: 1=Knows, 0=Doesn't know)
sample_data <- sample_data %>%
  mutate(know_hiv_test = ifelse(v756 == 1, 1, 0)) %>%
  filter(!is.na(know_hiv_test))

# Recode factors
sample_data <- sample_data %>%
  mutate(
    age_group = factor(v013, labels = c("15-19", "20-24", "25-29", "30-34", "35-39", "40-44", "45-49")),
    education = factor(v106, labels = c("No Education", "Primary", "Secondary", "Higher")),
    residence = factor(v025, labels = c("Urban", "Rural")),
    region = factor(v024),
    religion = factor(v130)
  )


#10-fold cross-validation
ctrl <- trainControl(method = "cv", number = 10, savePredictions = TRUE)
cv_model <- train(
  know_hiv_test ~ age_group + education + residence + region + religion,
  data = sample_data,
  method = "glm",
  family = "binomial",
  trControl = ctrl
)

# Output cross-validation results
cv_model$results
```

# 3) Consider data on CD4 count for 50 HIV/AIDS patients.258, 268, 264, 286, 270, 231, 253, 220, 240, 230, 240, 260, 240, 296, 266, 227, 264, 264,230, 254, 262, 267, 241, 231, 237, 244, 269, 247, 224, 246, 244, 219, 253, 238, 244, 243, 258, 222, 246, 254, 253, 229, 247, 221, 255, 253, 248, 241, 215, 220 Using the middle four digits of your ID in set.seed(), select a random sample of size 30 to answer the following question. Using bootstrapping estimate mean and median number of CD4count of the patients, include 95% CI for your estimates.

# Asnwer 

```{r results="hide"}
set.seed(9713)

# CD4 counts of the patients 
cd4_data <- c(258, 268, 264, 286, 270, 231, 253, 220, 240, 230, 240, 260, 240, 296, 266, 227, 264, 264,
              230, 254, 262, 267, 241, 231, 237, 244, 269, 247, 224, 246, 244, 219, 253, 238, 244, 243,
              258, 222, 246, 254, 253, 229, 247, 221, 255, 253, 248, 241, 215, 220)

# Set sample size and iterations
sample_size <- 30
iterations <- 10000  # number of bootstrap samples

# func bootstrapping to calculate mean and median
bootstrap_results <- function(data, iterations, sample_size) {
  # empty vectors for storing results 
  mean_values <- numeric(iterations)
  median_values <- numeric(iterations)
  
  # bootstrapping
  for (i in 1:iterations) {
    CD4_data_sample <- sample(data, size = sample_size, replace = TRUE)
    mean_values[i] <- mean(CD4_data_sample)
    median_values[i] <- median(CD4_data_sample)
  }
  
  # Calculate 95% confidence intervals
  mean_ci <- quantile(mean_values, c(0.025, 0.975))
  median_ci <- quantile(median_values, c(0.025, 0.975))
  
  # results
  return(list(mean = mean(mean_values), mean_ci = mean_ci,
              median = median(median_values), median_ci = median_ci))
}

# Perform bootstrapping
results <- bootstrap_results(cd4_data, iterations, sample_size)

# Output the results
cat("Estimated Mean: ", results$mean, "\n")
cat("95% CI for Mean: [", results$mean_ci[1], ", ", results$mean_ci[2], "]\n")
cat("Estimated Median: ", results$median, "\n")
cat("95% CI for Median: [", results$median_ci[1], ", ", results$median_ci[2], "]\n")

```
# The estimated mean CD4 count is 246.65, with a 95% confidence interval ranging from 240.47 to 253. The estimated median CD4 count is 246, with a 95% confidence interval between 240 and 254. This means we are 95% confident that the true mean and median CD4 counts fall within these intervals.

